\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{array}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  breaklines=true,
  frame=single,
  backgroundcolor=\color{gray!10},
  keywordstyle=\color{blue},
  commentstyle=\color{green!50!black},
  stringstyle=\color{red!60!black},
}

\title{Tetris OS Simulator: A Custom Operating System\\Module Demonstration Using Game Engine Architecture}
\author{
  \IEEEauthorblockN{Mehak Jain, Anuj Kumar Singh}
  \IEEEauthorblockA{Operating System}
}

\begin{document}

\maketitle

% ============================================================================
\begin{abstract}
This paper presents the design, implementation, and evaluation of \textbf{Tetris OS Simulator} --- a fully-functional Tetris game engine built from scratch in C without standard library dependencies for core logic. The project simulates seven fundamental operating system modules: Process Management, Memory Management, File Systems, I/O Management, Error Handling \& Security, Networking, and User Interface. The engine runs in two modes --- a terminal-based interface using ANSI escape codes and a WebSocket-backed browser client using React. Six custom libraries (\texttt{memory.c}, \texttt{string.c}, \texttt{math.c}, \texttt{screen.c}, \texttt{keyboard.c}, \texttt{sound.c}) replace all standard library functionality, while a hardware abstraction layer confines POSIX networking to dedicated I/O modules. The system demonstrates arena-based dynamic memory allocation, persistent file-based leaderboards, real-time WebSocket communication with automatic reconnection, and comprehensive error handling across all subsystems.
\end{abstract}

\begin{IEEEkeywords}
Operating Systems, Custom Libraries, Memory Management, WebSocket, Tetris, Process Scheduling, Hardware Abstraction
\end{IEEEkeywords}

% ============================================================================
\section{Introduction}

\subsection{Problem Statement}
Standard game engines rely heavily on standard library functions (\texttt{malloc}, \texttt{strlen}, \texttt{sin}, etc.) and opaque frameworks. This project challenges that paradigm by implementing a complete game system where all core logic is built from first principles --- custom arithmetic, custom string operations, custom memory allocation, and custom I/O drivers.

\subsection{Objectives}
\begin{enumerate}
  \item Implement six custom foundational libraries replacing \texttt{<string.h>}, \texttt{<math.h>}, and \texttt{malloc/free}.
  \item Build a complete Tetris game engine demonstrating all seven OS module categories.
  \item Provide dual I/O modes: terminal (ANSI) and browser (WebSocket).
  \item Ensure robust error handling, edge-case coverage, and graceful degradation.
  \item Maintain persistent state across sessions (file system module).
  \item Support real-time client-server communication (networking module).
\end{enumerate}

\subsection{Scope}
The project covers the complete lifecycle of an OS-like system: boot (memory initialization), process scheduling (game loop), I/O device management (keyboard/screen/speaker), persistent storage (leaderboard files), and graceful shutdown (memory cleanup, socket release).

% ============================================================================
\section{System Architecture}

The system follows a layered architecture where each custom library represents an OS module:

\begin{lstlisting}[language=C,caption={System Architecture Layers}]
Application Layer:  ws_tetris.c / main.c
Core Library Layer: string.c | math.c | memory.c
HAL Layer:          screen.c | keyboard.c | sound.c
Host OS / Hardware: POSIX sockets | Terminal | Audio
\end{lstlisting}

\textbf{Design Principle:} Game logic never includes networking headers, POSIX headers, or standard library functions beyond \texttt{<stdio.h>} and \texttt{<stdlib.h>}. All I/O flows through the Hardware Abstraction Layer (HAL).

% ============================================================================
\section{Methodology: Seven OS Modules}

% ---- Module 1 ----
\subsection{Module 1: Process Management}
Process management determines \textit{who gets to run, when, and where}. The game loop acts as a cooperative scheduler executing three processes each tick:

\begin{table}[h]
\centering
\caption{Process Schedule}
\begin{tabular}{lll}
\toprule
\textbf{Process} & \textbf{Priority} & \textbf{Frequency} \\
\midrule
Input Polling & Highest & Every tick (50ms) \\
Physics/Gravity & Medium & Level-dependent \\
Rendering & Lowest & On state change \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Speed Curve (NES-style):}
\begin{itemize}
  \item Level 1--8: \texttt{drop\_ticks = max(13 - level, 5)} --- 600ms to 250ms
  \item Level 9--13: \texttt{drop\_ticks = max(8 - (level-8)/2, 2)} --- 150ms to 100ms
  \item Level 14+: \texttt{drop\_ticks = 1} --- 50ms (maximum speed)
\end{itemize}

\textbf{Lock Delay:} A 10-tick (500ms) timer prevents instant locking. The timer resets on successful moves/rotations (capped at 15 resets).

\textbf{DAS/ARR:} Delayed Auto Shift (170ms delay) followed by Auto Repeat Rate (50ms interval) for smooth key repeat.

% ---- Module 2 ----
\subsection{Module 2: Memory Management}
Memory management determines \textit{who owns which memory and how much}.

\textbf{Design:} First-Fit Free-List Allocator with Block Splitting \& Coalescing.

\begin{lstlisting}[language=C,caption={BlockHeader Structure}]
typedef struct BlockHeader {
    int                  size;
    int                  is_free;
    struct BlockHeader  *next;
} BlockHeader;
\end{lstlisting}

\textbf{Operations:}
\begin{itemize}
  \item \texttt{memory\_init(65536)} --- acquires one 64KB slab via the single permitted \texttt{malloc()}.
  \item \texttt{t\_alloc(size)} --- first-fit search with 8-byte alignment; splits oversized blocks.
  \item \texttt{t\_dealloc(ptr)} --- validates pointer ownership, marks free, coalesces adjacent free blocks.
  \item \texttt{memory\_cleanup()} --- releases the slab back to the host OS.
\end{itemize}

\textbf{Safety guarantees:} Alignment to 8-byte boundaries, pointer validation via block-list walk, double-free detection, out-of-bounds rejection, and NULL returns on failure.

% ---- Module 3 ----
\subsection{Module 3: File System}
The file system provides \textit{persistent storage} for scores and leaderboard history.

\begin{itemize}
  \item \textbf{High Score File} (\texttt{highscore.txt}) --- terminal mode persists best score using \texttt{fopen/fgetc/fclose}.
  \item \textbf{Leaderboard File} (\texttt{leaderboard.txt}) --- WebSocket mode stores top 10 entries sorted by score.
  \item \textbf{Format:} \texttt{score|level|lines|name} (pipe-delimited, one entry per line).
  \item \textbf{Edge cases:} Missing file starts empty, malformed lines use safe defaults, write failures degrade gracefully.
\end{itemize}

% ---- Module 4 ----
\subsection{Module 4: I/O Management}
I/O management defines \textit{how this computer talks to the world} --- screen, keyboard, and speaker.

\begin{table}[h]
\centering
\caption{I/O Device Mapping}
\begin{tabular}{lll}
\toprule
\textbf{Device} & \textbf{Terminal} & \textbf{WebSocket} \\
\midrule
Keyboard & \texttt{getchar()} raw & \texttt{recv()} + decode \\
Screen & ANSI escapes & WebSocket frames \\
Speaker & WAV + \texttt{afplay} & Web Audio API \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Audio:} The terminal generates square-wave WAV files using pure integer arithmetic (no \texttt{<math.h>}), while the web client uses Web Audio API oscillators with an identical Korobeiniki melody.

% ---- Module 5 ----
\subsection{Module 5: Error Handling \& Security}
Error handling ensures \textit{graceful behaviour when things go wrong}.

\begin{table}[h]
\centering
\caption{Error Handling Matrix}
\begin{tabular}{ll}
\toprule
\textbf{Error Category} & \textbf{Handler} \\
\midrule
Division by zero & Returns 0 safely \\
Memory OOM & NULL-checked, graceful exit \\
Board out-of-bounds & \texttt{t\_in\_bounds()} guards \\
Buffer overflow & \texttt{t\_strncpy()} with max-length \\
Double-free & Silently ignored \\
Invalid pointer free & Validated against block list \\
Client disconnect & Re-accept without restart \\
Malformed JSON & Silently dropped \\
Handshake failure & Retry with next client \\
SIGINT & Terminal state restored \\
\bottomrule
\end{tabular}
\end{table}

% ---- Module 6 ----
\subsection{Module 6: Networking}
Networking defines \textit{how this computer talks to another computer}.

\textbf{Protocol:} WebSocket (RFC 6455) over TCP.

\textbf{Key features:}
\begin{itemize}
  \item Persistent server socket survives page refresh.
  \item Auto-reconnect client with 1.5s backoff.
  \item Full RFC 6455: SHA-1 handshake, masked frame decode, opcode handling.
  \item JSON state protocol: full game state serialized every dirty tick.
  \item Hardware abstraction: all socket headers confined to \texttt{screen.c} and \texttt{keyboard.c}.
\end{itemize}

% ---- Module 7 ----
\subsection{Module 7: User Interface}
The user interface defines \textit{how a user interacts with the game}.

\textbf{Terminal UI:} ANSI-colored 10x20 board with piece-type colors, ghost piece, next piece preview, HUD panel, game-over overlay, player name entry, sound effects, and background music.

\textbf{Web UI:} React + Vite with CRT scanline aesthetic, 4 selectable themes (Neon, Retro, Mono, Pastel), hold piece, ghost piece, 7-bag randomizer, SRS wall kicks, lock delay, DAS/ARR, combo system, level-up animations, leaderboard, touch controls, SFX toggle, and background music toggle.

% ============================================================================
\section{Implementation Details}

\subsection{Custom Library Pipeline}
\begin{lstlisting}[caption={Data Flow Pipeline}]
Input    -> keyboard.c  (captures raw input)
Parsing  -> string.c    (t_strcmp, t_itoa, t_strlen)
State    -> memory.c    (t_alloc for Game struct)
Logic    -> math.c      (t_mul, t_mod, t_div)
Output   -> screen.c    (ANSI / WebSocket frames)
Audio    -> sound.c     (WAV generation / Web Audio)
\end{lstlisting}

\subsection{Gameplay Mechanics}

\begin{table}[h]
\centering
\caption{Gameplay Feature Matrix}
\begin{tabular}{lcc}
\toprule
\textbf{Feature} & \textbf{Terminal} & \textbf{Web} \\
\midrule
7 Tetrominoes & \checkmark & \checkmark \\
Ghost Piece & \checkmark & \checkmark \\
Soft/Hard Drop & \checkmark & \checkmark \\
Level Progression & \checkmark & \checkmark \\
Sound Effects & \checkmark & \checkmark \\
Background Music & \checkmark & \checkmark \\
Player Name Entry & \checkmark & \checkmark \\
Score Persistence & \checkmark & \checkmark \\
Pause/Resume & \checkmark & \checkmark \\
SRS Wall Kicks & Basic & Full SRS \\
Hold Piece & \checkmark & \checkmark \\
7-Bag Randomizer & \checkmark & \checkmark \\
Lock Delay & \checkmark & \checkmark \\
DAS/ARR & $-$ & \checkmark \\
Combo System & \checkmark & \checkmark \\
Leaderboard & \checkmark & \checkmark \\
Theme Switcher & \checkmark & \checkmark \\
Touch Controls & $-$ & \checkmark \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Technology Stack}

\begin{table}[h]
\centering
\caption{Technology Components}
\begin{tabular}{ll}
\toprule
\textbf{Component} & \textbf{Technology} \\
\midrule
Backend & C (gcc -Wall -Wextra -O2) \\
Frontend & React 18 + Vite \\
Transport & WebSocket (RFC 6455) \\
Audio & WAV / Web Audio API \\
Persistence & Plain text files \\
Build & Make / npm \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================================
\section{Results and Testing}

\subsection{Compilation}
Both build targets compile with zero warnings under strict flags: \texttt{gcc -Wall -Wextra} (terminal) and \texttt{gcc -Wall -Wextra -O2} (backend).

\subsection{Edge Cases Verified}

\begin{table}[h]
\centering
\caption{Test Results}
\begin{tabular}{lcc}
\toprule
\textbf{Test Case} & \textbf{Expected} & \textbf{Result} \\
\midrule
Division by zero & Returns 0 & Pass \\
\texttt{t\_alloc(0)} & Returns NULL & Pass \\
\texttt{t\_dealloc(NULL)} & No-op & Pass \\
Double-free & Ignored & Pass \\
Page refresh & Re-accepts & Pass \\
Empty leaderboard & Starts empty & Pass \\
Name overflow (>20) & Truncated & Pass \\
Wall kick rotation & 5 tests & Pass \\
Hold twice per turn & Blocked & Pass \\
Malformed WebSocket & Dropped & Pass \\
Memory exhaustion & NULL logged & Pass \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Performance}
\begin{itemize}
  \item Game loop tick rate: 50ms (20 Hz physics)
  \item WebSocket frame size: 2--4 KB per state update
  \item Memory footprint: 64 KB arena (game state < 2 KB)
  \item Frontend bundle: < 100 KB (production)
\end{itemize}

% ============================================================================
\section{Future Scope}

\begin{enumerate}
  \item \textbf{Multiplayer Mode:} Extend WebSocket server for concurrent clients with garbage-line mechanics.
  \item \textbf{T-Spin Detection:} Implement 4-corner T-spin scoring (infrastructure already in place with \texttt{last\_was\_rotate} flag).
  \item \textbf{Sprint Mode:} Timed 40-line clear challenge demonstrating advanced process scheduling.
  \item \textbf{Replay System:} Record inputs with timestamps for playback (file system extension).
  \item \textbf{Cross-Platform Audio:} Replace macOS \texttt{afplay} with portable PCM output.
  \item \textbf{WebRTC Signaling:} Peer-to-peer multiplayer over the internet.
  \item \textbf{Persistent Sessions:} Token-based session tracking across server restarts.
\end{enumerate}

% ============================================================================
\section{Conclusion}

The Tetris OS Simulator successfully demonstrates all seven fundamental OS module categories through a practical, interactive application. By replacing all standard library functions with custom implementations, the project provides deep understanding of how operating systems manage processes, memory, I/O devices, file systems, and error conditions. The dual-mode architecture (terminal + WebSocket) showcases hardware abstraction principles, while the persistent leaderboard and real-time networking demonstrate file system and communication protocol implementations. Both build targets compile with zero warnings, and comprehensive edge-case testing confirms robust error handling across all subsystems.

% ============================================================================
\begin{thebibliography}{10}

\bibitem{c99}
ANSI/ISO 9899:1999, \textit{Programming Languages --- C}, International Organization for Standardization, 1999.

\bibitem{rfc6455}
I. Fette and A. Melnikov, ``The WebSocket Protocol,'' IETF RFC 6455, Dec. 2011. [Online]. Available: \url{https://tools.ietf.org/html/rfc6455}

\bibitem{sha1}
D. Eastlake and P. Jones, ``US Secure Hash Algorithm 1 (SHA-1),'' IETF RFC 3174, Sep. 2001.

\bibitem{guideline}
The Tetris Company, \textit{Tetris Guideline}, 2009. [Online]. Available: \url{https://tetris.wiki/Tetris_Guideline}

\bibitem{tanenbaum}
A. S. Tanenbaum and H. Bos, \textit{Modern Operating Systems}, 4th ed. Upper Saddle River, NJ: Pearson, 2014.

\bibitem{stallings}
W. Stallings, \textit{Operating Systems: Internals and Design Principles}, 9th ed. Upper Saddle River, NJ: Pearson, 2017.

\bibitem{kr}
B. W. Kernighan and D. M. Ritchie, \textit{The C Programming Language}, 2nd ed. Englewood Cliffs, NJ: Prentice Hall, 1988.

\bibitem{react}
React Team, ``React Documentation,'' Meta Open Source, 2024. [Online]. Available: \url{https://react.dev}

\bibitem{tls}
E. Rescorla, ``HTTP Over TLS,'' IETF RFC 2818, May 2000.

\bibitem{knuth}
D. E. Knuth, \textit{The Art of Computer Programming, Vol. 2: Seminumerical Algorithms}, 3rd ed. Boston, MA: Addison-Wesley, 1997.

\end{thebibliography}

\end{document}
